Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-06-06 paper_jasa.tex
not sure what i was chagning

The experiment is designed to study the ability of human observers to detect the effect of a single variable $X_2$ (corresponding to parameter $\beta_2$) in a two variable ($p=2$) regression model (Equation \ref{multi}). Data is simulated a range of values of $\beta_2~ (=0, 1, 3, 5, 7, 8, 10, 16)$, two different sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$). The range of $\beta_2$ values was chosen so that estimates of the power would produce reasonably continuous power curves, comparable to that calculated for the theoretical UMP test. Values were fixed for other regression parameters, $\beta_0 = 5$,  $\beta_1=15$, and the values for $X_1$ were randomly generated from a Poisson $(\lambda=30)$ distribution, which is almost Gaussian. %\green{Explain why this was used.} \red{There was not any real reason for this. It was just arbitrarily chosen to generate some data. When I first started to think of applying visual inference in regression setting, I was thinking of a dataset (I found in a book) of weight of babies by their gestational age (in week) where author wanted to test the difference of mean weight for gender.   Thus for our experiment I generated values similar to a variable like age. the discrete variable was thought similar to gender.}
Data sets were generated with parameter values shown in Table \ref{tbl:experiment_params}. Three replicates at each level were generated. These produced 60 different ``actual data sets'', and thus, 60 different lineups. To produce the lineups, the null model was fit to each actual data set, to obtain residuals and parameter estimates. The actual ydata plot was generated by making a side-by-side boxplot of the residuals (Table\ref{tbl:stat_multiple} case 3). The 19 null data sets were generated by simulating from  $N(0, {\hat{\sigma}}^2$, and these were plotted in the same way as side-by-side boxplots. The actual data plot was randomly placed among these null data plots to produce the lineup. Figure \ref{fig:test_category} is an example of one of these lineups. The actual data plot is number 11 which was generated for $n$=300, $\beta_2$=15 and $\sigma$=12.

%\green{I think it is plot11, not 13!!!!!!!! Please check!}

@@

 %Figure \ref{fig:test_category} is an example of one of these lineups. The visual test statistic is a side-by-side boxplot. The actual data plot is number 13 which was generated for $n$=300, $\beta_2$=15 and $\sigma$=12.

@@

{\bf Presenting lineups for evaluation} Each subject got at least 10 lineups for evaluation. For each combination of sample size ($n$) and error SD ($\sigma$) we define the difficulty of the lineup based on the five slope ($\beta$) parameters we choose as shown in table \ref{tbl:experiment_params}. For the smallest slope parameter the difficulty is 0 and for the largest parameter the difficulty is 4. The 10 lineups are chosen as follows;

@@

\begin{table}[hbtp]
\caption{Number of participants and their total feedbacks after applying five screening process on all three experimental data sets.}
\begin{center} {\small
\begin{tabular}{rrrrr|rrrr|rrrr}
  \hline 
Screening &  \multicolumn{4}{c} {Experiment 1}  & \multicolumn{4}{c} {Experiment 2}  & \multicolumn{4}{c} {Experiment 3} \\
 \cline{2-5}  \cline{6-9}   \cline{10-13}
criteria & total & male & female & feedbacks & total & male & female & feedbacks & total & male & female & feedbacks \\ 
  \hline
1 & 520 & 264 & 201 & 4914 & 390 & 203 & 182 & 4603 & 257 & 166 &  84 & 2830 \\ 
  2 & 465 & 264 & 201 & 4624 & 385 & 203 & 182 & 4558 & 250 & 166 &  84 & 2782 \\ 
  3 & 384 & 200 & 165 & 4187 & 384 & 199 & 181 & 4573 & 227 & 147 &  78 & 2648 \\ 
  4 & 318 & 168 & 135 & 3395 & 378 & 195 & 179 & 4492 & 205 & 133 &  72 & 2393 \\ 
  5 & 277 & 139 & 124 & 3008 & 374 & 193 & 177 & 4451 & 112 &  73 &  39 & 1410 \\  
   \hline
\end{tabular} }
\end{center}
\label{tbl:summary}
\end{table}

>>>>

The experiment is designed to study the ability of human observers to detect the effect of a single variable $X_2$ (corresponding to parameter $\beta_2$) in a two variable ($p=2$) regression model (Equation \ref{multi}). Data is simulated a range of values of $\beta_2~ (=0, 1, 3, 5, 7, 8, 10, 16)$, two different sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$). The range of $\beta_2$ values was chosen so that estimates of the power would produce reasonably continuous power curves, comparable to that calculated for the theoretical UMP test. Values were fixed for other regression parameters, $\beta_0 = 5$,  $\beta_1=15$, and the values for $X_1$ were randomly generated from a Poisson $(\lambda=30)$ distribution, which is almost Gaussian. Data sets were generated with parameter values shown in Table \ref{tbl:experiment_params}. Three replicates at each level were generated. These produced 60 different ``actual data sets'', and thus, 60 different lineups. To produce the lineups, the null model was fit to each actual data set, to obtain residuals and parameter estimates. The actual data plot was generated by making a side-by-side boxplot of the residuals (Table \ref{tbl:stat_multiple}, case 3). The 19 null data sets were generated by simulating from  $N(0, {\hat{\sigma}}^2)$, and these were plotted in the same way as side-by-side boxplots. The actual data plot was randomly placed among these null data plots to produce the lineup. Figure \ref{fig:test_category} is an example of one of these lineups. The actual data plot is number 11 which was generated for $n$=300, $\beta_2$=15 and $\sigma$=12.

@@

@@

{\bf Presenting lineups for evaluation} Each subject got at least 10 lineups for evaluation. For each combination of sample size ($n$) and error SD ($\sigma$) we define the difficulty of the lineup based on the five slope ($\beta$) parameters we choose as shown in table \ref{tbl:experiment_params}. For the smallest slope parameter the difficulty is 0 and for the largest parameter the difficulty is 4. The 10 lineups are choosen as follows;

@@

\begin{table}[hbtp]
\caption{Number of unique subjects and their total feedbacks after applying five screening process on all three experimental data sets.}
\begin{center}
\begin{tabular}{rrrrr|rrrr|rrrr}
  \hline
Screening &  \multicolumn{4}{c} {Experiment 1}  & \multicolumn{4}{c} {Experiment 2}  & \multicolumn{4}{c} {Experiment 3} \\
 \cline{2-5}  \cline{6-9}   \cline{10-13}
criteria & total & male & female & feedbacks & total & male & female & feedbacks & total & male & female & feedbacks \\ 
  \hline
1 & 520 & 264 & 201 & 4914 & 390 & 203 & 182 & 4603 & 257 & 166 &  84 & 2830 \\ 
  2 & 465 & 264 & 201 & 4624 & 385 & 203 & 182 & 4558 & 250 & 166 &  84 & 2782 \\ 
  3 & 384 & 200 & 165 & 4187 & 384 & 199 & 181 & 4573 & 227 & 147 &  78 & 2648 \\ 
  4 & 318 & 168 & 135 & 3395 & 378 & 195 & 179 & 4492 & 205 & 133 &  72 & 2393 \\ 
  5 & 277 & 139 & 124 & 3008 & 374 & 193 & 177 & 4451 & 112 &  73 &  39 & 1410 \\ 
   \hline
\end{tabular}
\end{center}
\label{tbl:summary}
\end{table}
